---
title: "Lab 2: Land Use Inventory Table"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: TRUE
fancy: true
geometry: margin=0.5in
documentclass: article
fontsize: 11pt
classoption: landscape
header-includes:
  \usepackage{setspace}
  \usepackage{titling}
  \setlength{\droptitle}{-10ex}
  \pretitle{\begin{flushleft}\Large\bfseries}
  \posttitle{\par\end{flushleft}}
  \preauthor{\begin{flushleft}\Large}
  \postauthor{\end{flushleft}}
  \predate{\begin{flushleft}}
  \postdate{\end{flushleft}}
---

\vspace{-2.2cm}
\raggedright
Jennah Gosciak
\newline
February 7th, 2022 \newline
Census Tracts Representing Borough Park: 232, 234, 472, 474, and 476
\vspace{0.5cm}

```{r, include = FALSE}
library(tidyverse)
library(tidycensus)
library(fs)
library(DBI)
library(dbplyr)
library(RPostgres)
library(knitr)
library(kableExtra)

con <- dbConnect(Postgres(),
  user = "furmandata",
  password = rstudioapi::askForPassword(),
  host = "fcdata.c7h93yqbqnvu.us-east-1.rds.amazonaws.com",
  port = 5432,
  dbname = "fcdata",
  bigint = "integer"
)

source("../99_common.R")
```
```{r, include = FALSE}
# Important link: http://maps.nyc.gov/doitt/nycitymap/
# need to collapse categories for this
lu_codes <- str_pad(c(0,1:11), width = 2, pad = "0")
lu_desc <- c("Residential (One- & Two-Family Buildings, Multifamily Walk-up Buildings, & Multifamily Elevator Buildings)",
  "One or Two Family Buildings",
  "Multi-Family Walk-Up Buildings",
  "Multi-Family Elevator Buildings",
  "Mixed Residential & Commerical Buildings",
  "Commercial & Office Buildings",
  "Industrial & Manufacturing Buildings",
  "Transportation & Utility",
  "Public Facilities & Institutions",
  "Open Space",
  "Parking Facilities",
  "Vacant Land"
)

sel_tracts <- c("232", "234",
                       "472",
                       "474",
                       "476")

df <- tbl(con, "mappluto_21v3") %>% 
  filter(ct2010 %in% sel_tracts) %>% 
  as.data.frame() 
df

df <- df %>% 
  filter(!is.na(landuse)) %>% 
  mutate(landuse_orig = landuse,
         landuse = case_when(landuse %in% c("01", "02", "03") ~ "00",
                             TRUE ~ landuse),
         land_use_desc = lu_desc[match(landuse, lu_codes)],
         land_use_desc = factor(land_use_desc, levels = lu_desc,
                                ordered = T))

# check recoding
df %>% 
  tablist_qc(landuse, land_use_desc, landuse_orig)
```
```{r, include = F}
create_lutable <- function(df) {
  df %>% 
    group_by(land_use_desc) %>%
    summarize(numlots = n_distinct(bbl),
              lotarea = sum(lotarea, na.rm = T),
              bldgarea = sum(bldgarea, na.rm = T)) %>% 
    ungroup() %>% 
    mutate(per_numlots = round(numlots * 100 / sum(numlots, na.rm = T)),
           per_lotarea = round(lotarea * 100 / sum(lotarea, na.rm = T)),
           per_bldgarea = round(bldgarea * 100 / sum(bldgarea, na.rm = T))) %>% 
    mutate(across(starts_with("per_"), ~str_c(., "%"))) %>% 
    select(land_use_desc,
           numlots, per_numlots,
           lotarea, per_lotarea,
           bldgarea, per_bldgarea)
}
```

\setlength{\tabcolsep}{6pt}
\renewcommand{\arraystretch}{1.2}
```{r, echo = FALSE}
df %>% 
  create_lutable() %>% 
  bind_rows(df %>%
              mutate(land_use_desc = "Total") %>% 
              create_lutable()) %>%
  rename_all(~c("Land Use", "Number of Lots", "Percentage of Total Number of Lots (%)",
               "Lot Area (sf)", "Percentage of Total Lot Area (sf)",
               "Building Floor Area (sf)", "Percentage of Total Building Floor Area (sf)")) %>%
  mutate(across(c("Number of Lots", "Lot Area (sf)",
                  "Building Floor Area (sf)"), ~scales::comma(., accuracy = 10))) %>% 
  kbl(align=c(rep('l',times=7))) %>% 
  kable_styling(position = "left",
                latex_options = c("hold_position")) %>%
  column_spec(1, width = "15em", bold = TRUE, border_left = TRUE) %>% 
  column_spec(2:6, width = "7em") %>% 
  column_spec(7, width = "7em", border_right = TRUE)  %>% 
  row_spec(0, bold = TRUE)
```
Data source: NYC Primary Land Use Tax Output (PLUTO)
\newpage
\onehalfspacing
The land use inventory table for Borough Park (represented by five selected census tracts) shows that the majority of all lots (81%) are residential. The most common land use categories by the number of lots are residential, mixed use, and public facilities and institutions. This is accurate to my perception of the neighborhood. Many buildings are either fully residential or have storefronts on the ground floor with residential apartments above. There are few buildings that are entirely commercial. Borough Park also has more than 300 religious institutions, one reason that there are so many lots under the category of public facilities and institutions (Beyer, 2010). The land use inventory table also shows a significant amount of vacant land (110 lots), which indicates areas that are either in the process of being redeveloped or areas that may be underutilized. Given the need for affordable housing, the Department of City Planning should evaluate whether these areas can be sources of new housing (Beyer, 2010).
\newline
\newline
The land use inventory table suggests that most buildings in Borough Park are low density. The ratio of total building floor area to lot area is 1.18 (almost 1), which suggests that for a number of buildings the lot area aligns with the building floor area. This is especially true for residential buildings since the ratio of building floor area to lot area is 1. For mixed use (residential and commercial), the ratio is 2.87, which indicates that mixed use buildings are higher density. Since vacant lots have a building floor area of 0, it may be slightly misleading to compare the total building floor area to the total lot area. For both lot area and building floor area, residential buildings comprise a smaller share than they did when analyzing land use by the number of lots. In contrast, public facilities and institutions comprise a larger share. This finding suggests that public facilities and institutions are likely larger building structures compared to residential buildings.
\newline
\singlespacing
Works Cited
\newline
Beyer, G. (2010, Oct. 8). Borough Park, Brooklyn. \textit{The New York Times.} [https://www.nytimes.com/2010/10/10/realestate/10living.html](https://www.nytimes.com/2010/10/10/realestate/10living.html).


