---
title: "Lab 3: Zoning Analysis"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: TRUE
fancy: true
geometry: margin=0.5in
documentclass: article
fontsize: 11pt
classoption: landscape
header-includes:
  \usepackage{setspace}
  \usepackage{titling}
  \setlength{\droptitle}{-10ex}
  \pretitle{\begin{flushleft}\Large\bfseries}
  \posttitle{\par\end{flushleft}}
  \preauthor{\begin{flushleft}\Large}
  \postauthor{\end{flushleft}}
  \predate{\begin{flushleft}}
  \postdate{\end{flushleft}}
---

\vspace{-2.2cm}
\raggedright
Jennah Gosciak
\newline
February 21st, 2022 \newline

```{r, include = FALSE}
library(tidyverse)
library(tidycensus)
library(fs)
library(DBI)
library(dbplyr)
library(RPostgres)
library(knitr)
library(kableExtra)

con <- dbConnect(Postgres(),
  user = "furmandata",
  password = rstudioapi::askForPassword(),
  host = "fcdata.c7h93yqbqnvu.us-east-1.rds.amazonaws.com",
  port = 5432,
  dbname = "fcdata",
  bigint = "integer"
)

source("../99_common.R")
```
```{r, include = FALSE}
# Important link: http://maps.nyc.gov/doitt/nycitymap/
sel_tracts <- c("366", "364",
                "362", "360.01",
                "360.02", "610.02",
                "610.04", "610.03")

df <- tbl(con, "mappluto_21v3") %>% 
  filter((ct2010 %in% sel_tracts) & borough == "BK") %>% 
  as.data.frame() 

stopifnot(length(unique(df$ct2010)) == 8)

df_sub <- df %>% 
  mutate(zonedist = str_sub(zonedist1, 1, 1),
         commarea = rowSums(select(., c("officearea", "retailarea", 
                                        "garagearea", "strgearea", 
                                        "otherarea")), na.rm = T)) %>% 
  rowwise() %>% 
  mutate(area_allowed = if_else(zonedist == "R",
                                residfar * lotarea,
                                commfar * lotarea),
         area_allowed_sp = if_else(is.na(spdist1), 0, 
                                   max(residfar, commfar, na.rm = T) * lotarea),
         max_area_allow = if_else(is.na(spdist1), area_allowed, area_allowed_sp),
         unutilized_res = if_else(is.na(spdist1), max(area_allowed - resarea, 
                                                      0, na.rm = T), 0),
         unutilized_comm_man = if_else(is.na(spdist1), 
                                       max(area_allowed - commarea - factryarea, 
                                           0, na.rm = T), 0),
         unutilized_sp = if_else(is.na(spdist1), 0, 
                                 max(area_allowed_sp - resarea - commarea - factryarea, 
                                     0, na.rm = T)),
         total_unutilized = case_when(!is.na(spdist1) ~ unutilized_sp,
                                      zonedist == "R" ~ unutilized_res,
                                      TRUE ~ unutilized_comm_man),
         utilized = bldgarea,
         lotunutilized = if_else(total_unutilized > 0 & !is.na(total_unutilized), "Yes", "No")) %>% 
  ungroup()

df_sub %>% 
  tablist_qc(zonedist, zonedist1)

df_sub %>% 
  select(c(zonedist, total_unutilized, utilized, max_area_allow))

df_sub %>% 
  select(c(unutilized_res, area_allowed, resarea)) %>% 
  mutate(check = area_allowed - resarea)

df_sub %>% 
  select(c(unutilized_comm_man, area_allowed, commarea, factryarea)) %>% 
  filter(commarea > 0)

df_sub %>% 
  select(c(area_allowed, residfar, lotarea, zonedist)) %>% 
  rowwise() %>% 
  mutate(check = residfar * lotarea)

df_sub %>% 
  filter(max_area_allow < utilized) %>% 
  mutate(check = residfar * lotarea) %>% 
  select("zonedist1", "bldgarea", "residfar", "lotarea", "max_area_allow", "unutilized_res",  "resarea", check, "yearbuilt", matches("area"))
```
```{r, include = FALSE}
create_lutable <- function(df) {
  df %>% 
    group_by(zonedist1) %>%
    summarize(numlots = n(),
              lotarea = sum(lotarea, na.rm = T),
              bldgarea = sum(bldgarea, na.rm = T)) %>% 
    ungroup() %>% 
    mutate(per_numlots = round(numlots * 100 / sum(numlots, na.rm = T)),
           per_lotarea = round(lotarea * 100 / sum(lotarea, na.rm = T)),
           per_bldgarea = round(bldgarea * 100 / sum(bldgarea, na.rm = T))) %>% 
    mutate(across(starts_with("per_"), ~str_c(., "%"))) %>% 
    select(zonedist1,
           numlots, per_numlots,
           lotarea, per_lotarea,
           bldgarea, per_bldgarea)
}
```

\setlength{\tabcolsep}{6pt}
\renewcommand{\arraystretch}{1.2}
\textbf{Zoning Inventory Analysis for Brighton Beach}
```{r, echo = FALSE, messages = FALSE, warning = FALSE}
df_sub %>% 
  create_lutable() %>% 
  bind_rows(df %>%
              mutate(zonedist1 = "Total") %>% 
              create_lutable()) %>%
  mutate(order = if_else(zonedist1 == "Total", 1, 0)) %>% 
  arrange(order, -numlots) %>% 
  select(-"order") %>% 
  rename_all(~c("Zoning Code", "Number of Lots", "Percentage of Total Number of Lots (%)",
               "Lot Area (sf)", "Percentage of Total Lot Area (sf)",
               "Building Floor Area (sf)", "Percentage of Total Building Floor Area (sf)")) %>%
  mutate(across(c("Number of Lots", "Lot Area (sf)",
                  "Building Floor Area (sf)"), ~scales::comma(.,
                                                              accuracy= 1))) %>% 
  kbl(align=c(rep('l',times=7))) %>% 
  kable_styling(position = "left",
                latex_options = c("hold_position")) %>%
  column_spec(1, width = "15em", bold = TRUE, border_left = TRUE) %>% 
  column_spec(2:6, width = "7em") %>% 
  column_spec(7, width = "7em", border_right = TRUE)  %>% 
  row_spec(0, bold = TRUE)
```
Data source: NYC Primary Land Use Tax Output (PLUTO)
\newpage
\setlength{\tabcolsep}{6pt}
\renewcommand{\arraystretch}{1.2}
\textbf{FAR Utilization Analysis for Brighton Beach}
```{r, echo = FALSE, warning = FALSE, message = FALSE}
create_lutable2 <- function(df) {
  df %>% 
    group_by(zonedist1) %>%
    summarize(across(c(total_unutilized,
                       utilized,
                       max_area_allow), sum, na.rm = T)) %>% 
    ungroup() %>% 
    mutate(per_unutil = round(total_unutilized * 100 / sum(total_unutilized, na.rm = T)),
           per_util = round(utilized * 100 / sum(utilized, na.rm = T)),
           per_max_allow = round(max_area_allow * 100 / sum(max_area_allow, na.rm = T))) %>% 
    mutate(across(starts_with("per_"), ~str_c(., "%"))) %>% 
    select(zonedist1, total_unutilized,
           per_unutil, utilized,
           per_util, max_area_allow,
           per_max_allow)
}

df_sub %>% 
  create_lutable2() %>% 
  bind_rows(df_sub %>%
              mutate(zonedist1 = "Total") %>% 
              create_lutable2()) %>%
  mutate(order = if_else(zonedist1 == "Total", 1, 0)) %>% 
  arrange(order, -total_unutilized) %>% 
  select(-"order") %>% 
  rename_all(~c("Zoning Code", "Total Unutilized FAR (sqft)", "Share of Total Unutilized FAR",
               "Total Utilized FAR (sqft)", "Share of Total Utilized FAR (sqft)",
               "Total Max Floor Area Allowed (sqft)", "Share of Total Max Floor Area Allowed")) %>%
  mutate(across(c("Total Unutilized FAR (sqft)",
                  "Total Utilized FAR (sqft)",
                  "Total Max Floor Area Allowed (sqft)"), ~scales::comma(., accuracy = 1))) %>% 
  kbl(align=c(rep('l',times=7))) %>% 
  kable_styling(position = "left",
                latex_options = c("hold_position")) %>%
  column_spec(1, width = "15em", bold = TRUE, border_left = TRUE) %>% 
  column_spec(2:6, width = "7em") %>% 
  column_spec(7, width = "7em", border_right = TRUE)  %>% 
  row_spec(0, bold = TRUE)
```

Data source: NYC Primary Land Use Tax Output (PLUTO)